Analytic Occultation Light Curves in Reflected Light

### 2020-04-16 tex/ms.tex
woot

\section{The Problem}
\label{sec:the-problem}
%
This paper closely follows the
notation and formalism introduced in \citet{Luger2019}. While we
include all of the relevant equations and definitions below,
the reader is encouraged to
review \citet{Luger2019} before proceeding.

\subsection{Review of the \starry algorithm in emitted light}
\label{sec:starry-review}
%
Without loss of generality, assume the body whose flux we wish to compute
has radius unity and sits at the origin of a right-handed Cartesian coordinate
system in some frame $\mathcal{F}_0$. In this frame,
the surface (emitted) intensity field of the body is described by a
vector $\mathbf{y}$ of coefficients in the spherical harmonic
basis $\by$:
%
\begin{align}
    \label{eq:by}
    \by & =
    \begin{pmatrix}
        Y_{0, 0}  &
        Y_{1, -1} & Y_{1, 0}  & Y_{1, 1} &
        Y_{2, -2} & Y_{2, -1} & Y_{2, 0} & Y_{2, 1} & Y_{2, 2} &
        \cdot\cdot\cdot
    \end{pmatrix}^\top
    \quad,
\end{align}
%
where the component at index $n$ is $Y_{l,m}$ with
%
\begin{align}
    \label{eq:l-m}
    l & = \floor*{\sqrt{n}}
    \nonumber               \\
    m & = n - l^2 - l
    \quad.
\end{align}
%
An observer views the body from a large distance in the sky frame
$\mathcal{F}$, in which the $x$-axis points to the right,
the $y$-axis points up, and the $z$-axis points out of the sky
toward the observer. Following \citet{Luger2019}, if
an occultor of radius $r_o$ is located at sky position $(x_o, y_o)$,
we compute the visible flux $f$ from
%
\begin{align}
    \label{eq:sTARRy}
    f = \sTe \mathbf{A} \mathbf{R}' \mathbf{R} \mathbf{y}
    \quad,
\end{align}
%
where, from right to left, $\mathbf{R} = \mathbf{R}(i, \lambda, \vartheta)$
is a Wigner rotation matrix that rotates $\bvec{y}$ from $\mathcal{F}_0$
to the sky frame $\mathcal{F}$
given the body's inclination $i$, obliquity
$\lambda$, and rotational phase $\vartheta$
\citep[Appendix C in][]{Luger2019},
%
$\mathbf{R}' = \mathbf{R}'(x_o, y_o)$ rotates the body on the plane
of the sky into the integration frame $\mathcal{F}'$, in which the
occultor lies along the $+y'$-axis,
%
$\mathbf{A}$
\citep[Equation~B13 in][]{Luger2019}
is the change-of-basis matrix from $\by$
to the \emph{Green's basis} $\bg$ in which the integrals are computed,
whose component at index $n$ is
%
\begin{align}
    \label{eq:bg}
    \tilde{g}_{n}(x, y) & =
    \begin{dcases}
        %
        \frac{\mu+2}{2}x^\frac{\mu}{2} y^\frac{\nu}{2}
         & \qquad \mu, \nu \, \text{even}
        \\[1em]
        %
        z(x, y)
         & \qquad \mu = \nu = 1
        \\[1em]
        %
        3x^{l-2}yz(x, y)
         & \qquad \nu \, \text{odd}, \,
        \mu = 1, \,
        \frac{\mu + \nu}{2} \, \text{even}
        \\[1em]
        %
        z(x, y)
        \bigg(
        -x^{l-3} + x^{l-1} + 4x^{l-3}y^2
        \bigg)
         & \qquad \nu \, \text{odd}, \,
        \mu = 1, \,
        \, \text{odd}
        \\[1em]
        %
        z(x, y)
        \bigg(
        \frac{\mu-3}{2} x^\frac{\mu-5}{2} y^\frac{\nu-1}{2}
        \ - \
        \frac{\mu-3}{2} x^\frac{\mu-5}{2} y^\frac{\nu+3}{2}
        \\
        \qquad\qquad \ - \
        \frac{\mu+3}{2} x^\frac{\mu-1}{2} y^\frac{\nu-1}{2}
        \bigg)
         & \qquad \text{otherwise}
        \quad,
    \end{dcases}
\end{align}
%
with
%
\begin{align}
    \label{eq:mu-nu}
    \mu & \equiv l - m
    \nonumber          \\
    \nu & \equiv l + m
\end{align}
%
and
%
\begin{align}
    \label{eq:z}
    z(x, y) \equiv \sqrt{1 - x^2 - y^2}
    \quad,
\end{align}
%
and $\sTe = \sTe(b_o, r_o)$ is the vector of solutions to the integral over
the projected visible disk of the body for each term in $\bg$
\citep[Equation~26 in][]{Luger2019}, with $b_o = \sqrt{x_o^2 + y_o^2}$.

If instead no occultor is present, we compute the total
visible flux $f_0$ from this body as
%
\begin{align}
    \label{eq:rTA1Ry}
    f_0 = \rTe \mathbf{A_1} \mathbf{R}'' \mathbf{R} \mathbf{y}
    \quad,
\end{align}
%
where, as before, $\mathbf{R} = \mathbf{R}(i, \lambda, \vartheta)$
rotates the body from $\mathcal{F}_0$
to the sky frame $\mathcal{F}$,
%
$\mathbf{R}''$ rotates the body on the plane
of the sky into the integration frame
$\mathcal{F}''$%
\footnote{%
    In \citet{Luger2019}, $\mathcal{F}'' = \mathcal{F}'$, so
    this rotation is trivial: $\mathbf{R}''$ is just the identity matrix.
},
%
$\mathbf{A_1}$
\citep[Equation~B11 in][]{Luger2019}
is the change-of-basis matrix from the spherical harmonic
basis $\by$ to the \emph{polynomial basis} $\bp$ in which the integrals
are computed, whose component at index $n$ is
%
\begin{align}
    \label{eq:bp}
    \tilde{p}_n(x, y) & =
    \begin{dcases}
        x^\frac{\mu}{2} y^\frac{\nu}{2}
         & \qquad \mu, \nu \, \text{even}
        \\[1em]
        x^\frac{\mu-1}{2} y^\frac{\nu-1}{2} z(x, y)
         & \qquad \text{otherwise}
        \quad,
    \end{dcases}
\end{align}
%
and $\rTe$ is the vector of solutions to the integral over
the projected visible disk of the body for each term in $\bp$
\citep[Equation~19 in][]{Luger2019}.

\subsection{Adapting the algorithm to the reflected light case}
\label{sec:starry-review}
%
In order to compute light curves in reflected light, we must make two
modifications to the \starry algorithm. First,
the expressions above assume that the coefficient vector
$\mathbf{y}$ describes the \emph{emissivity} of the body, which (in the
absence of limb darkening) is assumed to be Lambertian, i.e., all points on the
surface emit equally in all directions.
Here, we wish to derive the solution for the flux in the case of Lambertian
reflectance, in which case the vector $\mathbf{y}$ is taken to describe the
surface \emph{reflectivity} (or \emph{Bond albedo}).

Second, we must explicitly model the illumination of the body. We assume the
body is illuminated by a point-like source of unit luminosity, in which case
the radiance at any
point on the surface is proportional to the cosine of the angle between
the incident light and the surface normal. Points for which
this angle is greater than $\nicefrac{\pi}{2}$ are unilluminated and
therefore have radiance of zero.
%
If the point-like illumination source is placed at sky coordinates
$(x_s, y_s, z_s)$, the day/night terminator on the body is a half-ellipse
of semi-major axis unity that is fully described by its (signed) semi-minor
axis,
%
\begin{proof}{}
    \label{eq:b}
    b = -\frac{z_s}{r_s}
    \quad,
\end{proof}
%
where $r_s = \sqrt{x_s^2 + y_s^2 + z_s^2}$ is the distance to the source,
%
and the angle $\theta$ by which its semi-major axis is rotated away from the
$+x$-axis (a frame-dependent quantity, defined below).
Given this formulation, and assuming that
$r_s \gg 1$,
it is straightforward to show that the illumination
$\mathcal{I}$ at a point $(x, y)$ on the projected disk of the body is given
by the function
%
\begin{proof}{}
    \label{eq:illum}
    \mathcal{I}(b, \theta, r_s; x, y)&=
    \text{max}\bigg( 0, I(b, \theta, r_s; x, y) \bigg)
\end{proof}
%
where
%
\begin{proof}{}
    \label{eq:illum_poly}
    I(b, \theta, r_s; x, y) &= \frac{1}{4\pi r_s^2}
    \bigg(
    -b_c\sin\theta x + b_c\cos\theta y - bz(x, y)
    \bigg)
\end{proof}
%
with $b_c = \sqrt{1 - b^2}$ and $z(x, y) = \sqrt{1 - x^2 - y^2}$.

@@

@@

@@

We may now re-write Equations~(\ref{eq:sTARRy}) and (\ref{eq:rTA1Ry}) to
account for this illumination transformation. The flux during an occultation
is now given by
%
\begin{align}
    \label{eq:sTA2IA1RRy}
    f & =
    \sT(b, \theta', b_o, r_o)
    \mathbf{A_2}
    \mathbf{I}(b, \theta', r_s)
    \mathbf{A_1}
    \mathbf{R}'(x_o, y_o)
    \mathbf{R}(i, \lambda, \vartheta)
    \mathbf{y}
    \quad,
\end{align}
%
where
%
\begin{align}
    \label{eq:theta'}
    \theta' = \atantwo(x_o, y_o) - \atantwo(x_s, y_s)
\end{align}
%
is the angle of the terminator in the frame $\mathcal{F}'$.
Note that we made use of the fact that
$\mathbf{A} = \mathbf{A_2} \mathbf{A_1}$
\citep[Equation~14 in][]{Luger2019}, where $\mathbf{A_1}$ transforms from
the spherical harmonic basis $\by$ to the polynomial basis $\bp$, and
$\mathbf{A_2}$ transforms from $\bp$ to the Green's basis $\bg$.

Similarly, the flux when there is no occultation is now given by
%
\begin{align}
    \label{eq:rTIA1RRy}
    f_0 & =
    \rT(b)
    \mathbf{I}(b, \theta'', r_s)
    \mathbf{A_1}
    \mathbf{R}''(x_s, y_s)
    \mathbf{R}(i, \lambda, \vartheta)
    \mathbf{y}
    \quad,
\end{align}
%
where
%
\begin{align}
    \label{eq:theta''}
    \theta'' = 0
\end{align}
%
is the angle of the terminator in the frame $\mathcal{F}''$, by
construction. The transformation
$\mathbf{R}'' = \mathbf{R}''(x_s, y_s)$ rotates
the body
through an angle $\atantwo(x_s, y_s)$
so the semi-major axis of the terminator is aligned
with the $x''$-axis; as will become clear in \S\ref{sec:solution-no-occ} below,
this greatly simplifies the integration step.

@@

    \pmb{\rho}^\top(b, r_s) & \equiv
    \rT(b) \mathbf{I}(b, 0, r_s)

@@

trigonometric and algebraic functions of $b$. The component of $\pmb{\rho}^\top$

@@

    \rho_n(b, r_s) & =

@@

        J_{\frac{\mu}{2}, \frac{\nu + 2}{2}} -
        b H_\frac{\nu}{2}(b) K_{\frac{\mu}{2}, \frac{\nu}{2}}

@@

        H_{\frac{\nu + 1}{2}}(b) K_{\frac{\mu - 1}{2}, \frac{\nu + 1}{2}}

@@

        J_{\frac{\mu - 1}{2}, \frac{\nu - 1}{2}} -
        J_{\frac{\mu + 3}{2}, \frac{\nu - 1}{2}}

@@

        J_{\frac{\mu - 1}{2}, \frac{\nu + 3}{2}}

@@

where we make use of the indices
    \label{eq:l-m-mu-nu}
    \mu & = l - m
    \nu & = l + m
    l   & = \floor*{\sqrt{n}}
    m   & = n - l^2 - l
as in \citet{Luger2019} and we define the helper functions

@@

    H_{j}(b) &= \int_b^1 a^j \sqrt{1 - a^2} \dd a
    J_{i,j} &=

@@

    K_{i,j} &=

@@

@@

        H_{0}(b) & = \frac{\arccos(b) - bb_c}{2}
        H_{1}(b) & = \frac{b_c^3}{3}

@@

        J_{0,0} & = \pi
        J_{0,1} & = \frac{4}{3}

@@

        K_{0,0} &= \frac{4}{3}
        K_{0,1} &= \frac{3\pi}{8}

@@

    H_{j}(b) &= \frac{b^n b_c^3 + (j - 1) H_{j - 2}(b)}{j + 2}
    J_{0,j} &= \left(\frac{j - 1}{j + 2}\right) J_{0,j-2}
    K_{0,j} &= \left(\frac{j + 2}{j + 3}\right) K_{0,j-2}
    J_{i,j} &= \left(\frac{i - 1}{i + j + 2}\right) J_{i-2,j}
    K_{i,j} &= \left(\frac{i - 1}{i + j + 3}\right) K_{i-2,j}

@@

Once $\pmb{\rho}^\top$ is known, the observed flux is computed from

@@

    \pmb{\rho}^\top(b, r_s)

@@

    \mathbf{R}(i, \lambda, \vartheta)

@@

    -\pmb{\rho}^\top(-b, r_s)

@@

The key to identifying the case corresponding to a given configuration is
to determine whether or not the limb of the occultor intersects the terminator
of the body, and if so, the points of intersection.
While we perform the integration in frame $\mathcal{F}'$, finding the points
of intersection with the terminator is easier if we temporarily switch to
the frame $\mathcal{F}''$ (via Equation~\ref{eq:R''}) in which the terminator
is parallel to the $x''$-axis.
In this frame, the equations defining the terminator and the limb of the
occultor are, respectively,
%
\begin{proof}{}
    y''_1(x'') & = b \sqrt{1 - x''^2}
    \nonumber                                               \\
    y''_2(x'') & = y''_o \pm \sqrt{r_o^2 - (x'' - x''_o)^2}
\end{proof}
%
where
%
\begin{proof}{}
    x''_o = b_o\sin\theta'
    \nonumber \\
    y''_o = b_o\cos\theta'
\end{proof}
%
are the coordinates of the occultor in $\mathcal{F}''$.
%
We wish to find the vector of $N$ points
$\mathbf{x''} = \left(x_0, x_1, {\cdot\cdot\cdot}, x_{N-1}\right)^\top$
for which
$y''_1(x_n'') - y''_2(x_n'') = 0$. Following \citet{Luger2017}, we may
express this condition as the quartic equation
%
\begin{proof}{}
    \label{eq:quartic}
    A {x''}^4 + B {x''}^3 + C {x''}^2 + D {x''} + E = 0
\end{proof}
%
with coefficients
%
\begin{proof}{}
    \label{eq:quartic-coeffs}
    A &= (1 - b^2)^2
    \nonumber \\
    B &= -4 x''_o (1 - b^2)
    \nonumber \\
    C &= -2 \bigg(
    b^4
    + r_o^2
    - 3 {x''_o}^2
    - {y''_o}^2
    - b^2 \big(1 + r_o^2 - {x''_o}^2 + {y''_o}^2\big)
    \bigg)
    \nonumber \\
    D &= -4 x''_o (b^2 - r_o^2 + {x''_o}^2 + {y''_o}^2)
    \nonumber \\
    E &=
    b^4
    - 2 b^2 \big(r_o^2 - {x''_o}^2 + {y''_o}^2\big)
    + \big(r_o^2 - {x''_o}^2 - {y''_o}^2\big)^2
    \quad.
\end{proof}
%
Although closed-form solutions to quartic equations exist
\citep[see, e.g.,][who solve for the area of overlap between two ellipses
    analytically]{Hughes2011}, they are prone to significant numerical
instabilities. Instead, we solve for the roots of the quartic
numerically by casting it
as an eigenvalue problem \citep[e.g.,][]{Edelman1995} and polish the
results with a few iterations of Newton's method. We find that this is
reasonably computationally efficient and
yields roots with precision within a couple orders of magnitude of machine
epsilon (see \S\ref{sec:performance}).

@@

\subsection{Cases 1--5}
\label{sec:cases-easy}
%
Cases 1--5 (see Figure~\ref{fig:cases}) involve configurations in which the
occultor does not intersect with
the terminator of the occulted body, and are therefore fairly
straightforward to solve. In particular, we can use the original emitted
light solution from \citet{Luger2019}, provided we weight the map by our
polynomial illumination function:
%
\begin{proof}{}
    \label{eq:fI}
    f_I &=
    \sTe(b_o, r_o)
    \mathbf{A_2}
    \mathbf{I}(b, \theta', r_s)
    \mathbf{A_1}
    \mathbf{R}'(x_o, y_o)
    \mathbf{R}(i, \lambda, \vartheta)
    \mathbf{y}
    \quad,
\end{proof}
%
where $\sTe(b_o, r_o)$ is the emitted light solution vector
\citep[Equation~26 in][]{Luger2019}. The flux $f_I$ is the flux one would
measure from a body whose surface map is weighted by the illumination function
$\mathbf{I}(b, \theta', r_s)$ during an occultation. Note that this is not
necessarily the \emph{observed} flux, since this may include the unphysical
negative contribution from the nightside. We must compute the actual
observed flux on a case-by-case basis.

@@

@@

The components of the vector $\viH$ are given by the integral

@@

Given $\vlambda$, the solution to Equation~(\ref{eq:qT}) is straightforward:
%
\begin{proof}{}
    \mathbb{q}_n &=
    \begin{cases}
        \iH_{\frac{\mu + 4}{2}, \frac{\nu}{2}}(\vlambda)
                                  & \qquad \mu, \nu, \, \text{even}
        \\[1em]
        \frac{1}{3}\Delta\vlambda & \qquad \mu = \nu = 1
        \\[1em]
        0                         & \qquad \text{otherwise}
        \quad,
    \end{cases}
\end{proof}
%
where the vector $\viH$ is given by Equation~(\ref{eq:H}).

>>>>

\section{The Problem}
\label{sec:the-problem}
%
This paper closely follows the
notation and formalism introduced in \citet{Luger2019}. While we
include all of the relevant equations and definitions below,
the reader is encouraged to
review \citet{Luger2019} before proceeding.
To improve the readability of this paper,
Tables~\ref{tab:symbols}--\ref{tab:matrices} at the end list the principal
symbols and quantities used throughout the text, with links to the
sections and equations in which they are defined.

\subsection{Review of the \starry algorithm in emitted light}
\label{sec:starry-review}
%
Without loss of generality, assume the body whose flux we wish to compute
has radius unity and sits at the origin of a right-handed Cartesian coordinate
system in some frame $\mathcal{F}_0$. In this frame,
the surface (emitted) intensity field of the body is described by a
vector $\mathbf{y}$ of coefficients in the spherical harmonic
basis $\by$:
%
\begin{align}
    \label{eq:by}
    \by & =
    \begin{pmatrix}
        Y_{0, 0}  &
        Y_{1, -1} & Y_{1, 0}  & Y_{1, 1} &
        Y_{2, -2} & Y_{2, -1} & Y_{2, 0} & Y_{2, 1} & Y_{2, 2} &
        \cdot\cdot\cdot
    \end{pmatrix}^\top
    \quad,
\end{align}
%
where the component at index $n$ is $Y_{l,m}$ with
%
\begin{align}
    \label{eq:l-m}
    l & = \floor*{\sqrt{n}}
    \nonumber               \\
    m & = n - l^2 - l
    \quad.
\end{align}
%
An observer views the body from a large distance in the sky frame
$\mathcal{F}$, in which the $x$-axis points to the right,
the $y$-axis points up, and the $z$-axis points out of the sky
toward the observer. Following \citet{Luger2019}, if
an occultor of radius $r_o$ is located at sky position $(x_o, y_o)$,
we compute the visible flux $f$ from
%
\begin{align}
    \label{eq:sTARRy}
    f = \sTe \mathbf{A} \mathbf{R}' \mathbf{R} \mathbf{y}
    \quad,
\end{align}
%
where, from right to left, $\mathbf{R} = \mathbf{R}(\text{I}, \Lambda, \Theta)$
is a Wigner rotation matrix that rotates $\bvec{y}$ from $\mathcal{F}_0$
to the sky frame $\mathcal{F}$
given the body's inclination $\text{I}$, obliquity
$\Lambda$, and rotational phase $\Theta$
\citep[Appendix C in][]{Luger2019},
%
$\mathbf{R}' = \mathbf{R}'(x_o, y_o)$ rotates the body on the plane
of the sky into the integration frame $\mathcal{F}'$, in which the
occultor lies along the $+y'$-axis,
%
$\mathbf{A}$
\citep[Equation~B13 in][]{Luger2019}
is the change-of-basis matrix from $\by$
to the \emph{Green's basis} $\bg$ in which the integrals are computed,
whose component at index $n$ is
%
\begin{align}
    \label{eq:bg}
    \tilde{g}_{n}(x, y) & =
    \begin{dcases}
        %
        \frac{\mu+2}{2}x^\frac{\mu}{2} y^\frac{\nu}{2}
         & \qquad \mu, \nu \, \text{even}
        \\[1em]
        %
        z(x, y)
         & \qquad \mu = \nu = 1
        \\[1em]
        %
        3x^{l-2}yz(x, y)
         & \qquad \nu \, \text{odd}, \,
        \mu = 1, \,
        \frac{\mu + \nu}{2} \, \text{even}
        \\[1em]
        %
        z(x, y)
        \bigg(
        -x^{l-3} + x^{l-1} + 4x^{l-3}y^2
        \bigg)
         & \qquad \nu \, \text{odd}, \,
        \mu = 1, \,
        \, \text{odd}
        \\[1em]
        %
        z(x, y)
        \bigg(
        \frac{\mu-3}{2} x^\frac{\mu-5}{2} y^\frac{\nu-1}{2}
        \ - \
        \frac{\mu-3}{2} x^\frac{\mu-5}{2} y^\frac{\nu+3}{2}
        \\
        \qquad\qquad \ - \
        \frac{\mu+3}{2} x^\frac{\mu-1}{2} y^\frac{\nu-1}{2}
        \bigg)
         & \qquad \text{otherwise}
        \quad,
    \end{dcases}
\end{align}
%
with
%
\begin{align}
    \label{eq:mu-nu}
    \mu & \equiv l - m
    \nonumber          \\
    \nu & \equiv l + m
\end{align}
%
and
%
\begin{align}
    \label{eq:z}
    z(x, y) \equiv \sqrt{1 - x^2 - y^2}
    \quad,
\end{align}
%
and $\sTe = \sTe(b_o, r_o)$ is the vector of solutions to the integral over
the projected visible disk of the body for each term in $\bg$
\citep[Equation~26 in][]{Luger2019}, with $b_o = \sqrt{x_o^2 + y_o^2}$.

If instead no occultor is present, we compute the total
visible flux $f_0$ from this body as
%
\begin{align}
    \label{eq:rTA1Ry}
    f_0 = \rTe \mathbf{A_1} \mathbf{R}'' \mathbf{R} \mathbf{y}
    \quad,
\end{align}
%
where, as before, $\mathbf{R} = \mathbf{R}(\text{I}, \Lambda, \Theta)$
rotates the body from $\mathcal{F}_0$
to the sky frame $\mathcal{F}$,
%
$\mathbf{R}''$ rotates the body on the plane
of the sky into the integration frame
$\mathcal{F}''$%
\footnote{%
    In \citet{Luger2019}, $\mathcal{F}'' = \mathcal{F}'$, so
    this rotation is trivial: $\mathbf{R}''$ is just the identity matrix.
},
%
$\mathbf{A_1}$
\citep[Equation~B11 in][]{Luger2019}
is the change-of-basis matrix from the spherical harmonic
basis $\by$ to the \emph{polynomial basis} $\bp$ in which the integrals
are computed, whose component at index $n$ is
%
\begin{align}
    \label{eq:bp}
    \tilde{p}_n(x, y) & =
    \begin{dcases}
        x^\frac{\mu}{2} y^\frac{\nu}{2}
         & \qquad \mu, \nu \, \text{even}
        \\[1em]
        x^\frac{\mu-1}{2} y^\frac{\nu-1}{2} z(x, y)
         & \qquad \text{otherwise}
        \quad,
    \end{dcases}
\end{align}
%
and $\rTe$ is the vector of solutions to the integral over
the projected visible disk of the body for each term in $\bp$
\citep[Equation~19 in][]{Luger2019}.

\subsection{Adapting the algorithm to the reflected light case}
\label{sec:adapting-starry}
%
In order to compute light curves in reflected light, we must make two
modifications to the \starry algorithm. First,
the expressions above assume that the coefficient vector
$\mathbf{y}$ describes the \emph{emissivity} of the body, which (in the
absence of limb darkening) is assumed to be Lambertian, i.e., all points on the
surface emit equally in all directions.
Here, we wish to derive the solution for the flux in the case of Lambertian
reflectance, in which case the vector $\mathbf{y}$ is taken to describe the
surface \emph{reflectivity} (or \emph{Bond albedo}).

Second, we must explicitly model the illumination of the body. We assume the
body is illuminated by a point-like source of unit luminosity, in which case
the radiance at any
point on the surface is proportional to the cosine of the angle between
the incident light and the surface normal. Points for which
this angle is greater than $\nicefrac{\pi}{2}$ are unilluminated and
therefore have radiance of zero.
%
If the point-like illumination source is placed at sky coordinates
$(x_s, y_s, z_s)$, the day/night terminator on the body is a half-ellipse
of semi-major axis unity that is fully described by its (signed) semi-minor
axis,
%
\begin{proof}{}
    \label{eq:b}
    b = -\frac{z_s}{r_s}
    \quad,
\end{proof}
%
where $r_s = \sqrt{x_s^2 + y_s^2 + z_s^2}$ is the distance to the source,
%
and the angle $\theta$ by which its semi-major axis is rotated away from the
$+x$-axis (a frame-dependent quantity, defined below).
Given this formulation, and assuming that
$r_s \gg 1$,
it is straightforward to show that the illumination
$\mathcal{I}$ at a point $(x, y)$ on the projected disk of the body is given
by the function
%
\begin{proof}{}
    \label{eq:illum}
    \mathcal{I}(b, \theta, r_s; x, y)&=
    \text{max}\bigg( 0, I(b, \theta, r_s; x, y) \bigg)
\end{proof}
%
where
%
\begin{proof}{}
    \label{eq:illum_poly}
    I(b, \theta, r_s; x, y) &= \frac{1}{4\pi r_s^2}
    \bigg(
    -b_c\sin\theta x + b_c\cos\theta y - bz(x, y)
    \bigg)
\end{proof}
%
with $b_c \equiv \sqrt{1 - b^2}$ and $z(x, y) = \sqrt{1 - x^2 - y^2}$.

@@

    \label{eq:ivec}

@@

    \label{eq:Imat}

@@

We may now re-write Equations~(\ref{eq:sTARRy}) and (\ref{eq:rTA1Ry}) to
account for this illumination transformation. The flux during an occultation
is now given by
%
\begin{align}
    \label{eq:sTA2IA1RRy}
    f & =
    \sT(b, \theta', b_o, r_o)
    \mathbf{A_2}
    \mathbf{I}(b, \theta', r_s)
    \mathbf{A_1}
    \mathbf{R}'(x_o, y_o)
    \mathbf{R}(\text{I}, \Lambda, \Theta)
    \mathbf{y}
    \quad,
\end{align}
%
where
%
\begin{align}
    \label{eq:theta'}
    \theta' = \atantwo(x_o, y_o) - \atantwo(x_s, y_s)
\end{align}
%
is the angle of the terminator in the frame $\mathcal{F}'$.
Note that we made use of the fact that
$\mathbf{A} = \mathbf{A_2} \mathbf{A_1}$
\citep[Equation~14 in][]{Luger2019}, where $\mathbf{A_1}$ transforms from
the spherical harmonic basis $\by$ to the polynomial basis $\bp$, and
$\mathbf{A_2}$ transforms from $\bp$ to the Green's basis $\bg$.

Similarly, the flux when there is no occultation is now given by
%
\begin{align}
    \label{eq:rTIA1RRy}
    f_0 & =
    \rT(b)
    \mathbf{I}(b, \theta'', r_s)
    \mathbf{A_1}
    \mathbf{R}''(x_s, y_s)
    \mathbf{R}(\text{I}, \Lambda, \Theta)
    \mathbf{y}
    \quad,
\end{align}
%
where
%
\begin{align}
    \label{eq:theta''}
    \theta'' = 0
\end{align}
%
is the angle of the terminator in the frame $\mathcal{F}''$, by
construction. The transformation
$\mathbf{R}'' = \mathbf{R}''(x_s, y_s)$ rotates
the body
through an angle $\atantwo(x_s, y_s)$
so the semi-major axis of the terminator is aligned
with the $x''$-axis; as will become clear in \S\ref{sec:solution-no-occ} below,
this greatly simplifies the integration step.

@@

    \overset{\quad\pmb{\wedge}}{\mathbbb{r}}{^\top}(b, r_s) & \equiv
    \rT(b) \mathbf{I}(b, \theta'', r_s)

@@

trigonometric and algebraic functions of $b$.
The component of $\overset{\quad\pmb{\wedge}}{\mathbbb{r}}{^\top}$

@@

    \overset{\quad\wedge}{\mathbb{r}}_n(b, r_s) & =

@@

        \mathbb{L}_{\frac{\mu}{2}, \frac{\nu + 2}{2}} -
        b \mathbb{k}_\frac{\nu}{2}(b) \mathbb{M}_{\frac{\mu}{2}, \frac{\nu}{2}}

@@

        \mathbb{k}_{\frac{\nu + 1}{2}}(b) \mathbb{M}_{\frac{\mu - 1}{2}, \frac{\nu + 1}{2}}

@@

        \mathbb{L}_{\frac{\mu - 1}{2}, \frac{\nu - 1}{2}} -
        \mathbb{L}_{\frac{\mu + 3}{2}, \frac{\nu - 1}{2}}

@@

        \mathbb{L}_{\frac{\mu - 1}{2}, \frac{\nu + 3}{2}}

@@

where the components of $\mathbbb{k}$, $\mathbbb{L}$, and $\mathbbb{M}$
are given by

@@

    \mathbb{k}_{j}(b) &= \int_b^1 a^j \sqrt{1 - a^2} \dd a
    \mathbb{L}_{i,j} &=

@@

    \mathbb{M}_{i,j} &=

@@

where $\Gamma$ is the gamma function.

@@

        \mathbb{k}_{0}(b) & = \frac{\arccos(b) - bb_c}{2}
        \mathbb{k}_{1}(b) & = \frac{b_c^3}{3}

@@

        \mathbb{L}_{0,0} & = \pi
        \mathbb{L}_{0,1} & = \frac{4}{3}

@@

        \mathbb{M}_{0,0} &= \frac{4}{3}
        \mathbb{M}_{0,1} &= \frac{3\pi}{8}

@@

    \mathbb{k}_{j}(b) &= \frac{b^n b_c^3 + (j - 1) \mathbb{k}_{j - 2}(b)}{j + 2}
    \mathbb{L}_{0,j} &= \left(\frac{j - 1}{j + 2}\right) \mathbb{L}_{0,j-2}
    \mathbb{M}_{0,j} &= \left(\frac{j + 2}{j + 3}\right) \mathbb{M}_{0,j-2}
    \mathbb{L}_{i,j} &= \left(\frac{i - 1}{i + j + 2}\right) \mathbb{L}_{i-2,j}
    \mathbb{M}_{i,j} &= \left(\frac{i - 1}{i + j + 3}\right) \mathbb{M}_{i-2,j}

@@

Once $\overset{\quad\pmb{\wedge}}{\mathbbb{r}}{^\top}$ is known,
the observed flux is computed from

@@

    \overset{\quad\pmb{\wedge}}{\mathbbb{r}}{^\top}(b, r_s)

@@

    \mathbf{R}(\text{I}, \Lambda, \Theta)

@@

    -\overset{\quad\pmb{\wedge}}{\mathbbb{r}}{^\top}(-b, r_s)

@@

The key to identifying the case corresponding to a given configuration is
to determine whether or not the limb of the occultor intersects the terminator
of the body, and if so, the points of intersection.
While we perform the integration in frame $\mathcal{F}'$, finding the points
of intersection with the terminator is easier if we temporarily switch to
the frame $\mathcal{F}''$, in which the terminator
is parallel to the $x''$-axis.
In this frame, the equations defining the terminator and the limb of the
occultor are, respectively,
%
\begin{proof}{}
    y''_1(x'') & = b \sqrt{1 - x''^2}
    \nonumber                                               \\
    y''_2(x'') & = y''_o \pm \sqrt{r_o^2 - (x'' - x''_o)^2}
\end{proof}
%
where
%
\begin{proof}{}
    x''_o = b_o\sin\theta'
    \nonumber \\
    y''_o = b_o\cos\theta'
\end{proof}
%
are the coordinates of the occultor in $\mathcal{F}''$.
%
We wish to find the vector of $N$ points
$\mathbf{x''} = \left(x_0, x_1, {\cdot\cdot\cdot}, x_{N-1}\right)^\top$
for which
$y''_1(x_n'') - y''_2(x_n'') = 0$. Following \citet{Luger2017}, we may
express this condition as the quartic equation
%
\begin{proof}{}
    \label{eq:quartic}
    A {x''}^4 + B {x''}^3 + C {x''}^2 + D {x''} + E = 0
\end{proof}
%
with coefficients
%
\begin{proof}{}
    \label{eq:quartic-coeffs}
    A &= (1 - b^2)^2
    \nonumber \\
    B &= -4 x''_o (1 - b^2)
    \nonumber \\
    C &= -2 \bigg(
    b^4
    + r_o^2
    - 3 {x''_o}^2
    - {y''_o}^2
    - b^2 \big(1 + r_o^2 - {x''_o}^2 + {y''_o}^2\big)
    \bigg)
    \nonumber \\
    D &= -4 x''_o (b^2 - r_o^2 + {x''_o}^2 + {y''_o}^2)
    \nonumber \\
    E &=
    b^4
    - 2 b^2 \big(r_o^2 - {x''_o}^2 + {y''_o}^2\big)
    + \big(r_o^2 - {x''_o}^2 - {y''_o}^2\big)^2
    \quad.
\end{proof}
%
Although closed-form solutions to quartic equations exist
\citep[see, e.g.,][who solve for the area of overlap between two ellipses
    analytically]{Hughes2011}, they are prone to significant numerical
instabilities. Instead, we solve for the roots of the quartic
numerically by casting it
as an eigenvalue problem \citep[e.g.,][]{Edelman1995} and polish the
results with a few iterations of Newton's method. We find that this is
reasonably computationally efficient and
yields roots with precision within a couple orders of magnitude of machine
epsilon (see \S\ref{sec:performance}).

@@

\subsection{Cases 1--5}
\label{sec:cases-easy}
%
Cases 1--5 (see Figure~\ref{fig:cases}) involve configurations in which the
occultor does not intersect with
the terminator of the occulted body, and are therefore fairly
straightforward to solve. In particular, we can use the original emitted
light solution from \citet{Luger2019}, provided we weight the map by our
polynomial illumination function:
%
\begin{proof}{}
    \label{eq:fI}
    f_I &=
    \sTe(b_o, r_o)
    \mathbf{A_2}
    \mathbf{I}(b, \theta', r_s)
    \mathbf{A_1}
    \mathbf{R}'(x_o, y_o)
    \mathbf{R}(\text{I}, \Lambda, \Theta)
    \mathbf{y}
    \quad,
\end{proof}
%
where $\sTe(b_o, r_o)$ is the emitted light solution vector
\citep[Equation~26 in][]{Luger2019}. The flux $f_I$ is the flux one would
measure from a body whose surface map is weighted by the illumination function
$\mathbf{I}(b, \theta', r_s)$ during an occultation. Note that this is not
necessarily the \emph{observed} flux, since this may include the unphysical
negative contribution from the nightside. We must compute the actual
observed flux on a case-by-case basis.

@@

    \label{eq:V}

@@

The components of the matrix $\viH$ are given by the integral

@@

Given $\vlambda$, the solution to Equation~(\ref{eq:qT}) is straightforward:
%
\begin{proof}{}
    \mathbb{q}_n &=
    \begin{cases}
        \iH_{\frac{\mu + 4}{2}, \frac{\nu}{2}}(\vlambda)
                                  & \qquad \mu, \nu, \, \text{even}
        \\[1em]
        \frac{1}{3}\Delta\vlambda & \qquad \mu = \nu = 1
        \\[1em]
        0                         & \qquad \text{otherwise}
        \quad,
    \end{cases}
\end{proof}
%
where the matrix $\viH$ is given by Equation~(\ref{eq:H}).
